---
title: "Experimental cluster-permutation prototype"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experimental cluster-permutation prototype}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 4.8
)
```

# Purpose

This article demonstrates the experimental cluster-permutation prototype in gpbiometrics.

The implementation is deliberately narrow. It currently supports two-condition, within-subject, one-dimensional time-course comparisons using participant-level condition time courses.

# Interpretation warning

A significant cluster indicates evidence against the global null of no condition difference anywhere in the tested time range under the permutation scheme.

It does not identify a precise onset, offset, latency, or physiological event boundary. Cluster timing should be treated as descriptive.

# Synthetic example

```{r}
library(gpbiometrics)

set.seed(123)

dat <- expand.grid(
  participant = sprintf("P%02d", 1:12),
  condition = c("A", "B"),
  time = 1:60,
  KEEP.OUT.ATTRS = FALSE,
  stringsAsFactors = FALSE
)

subject_shift <- rnorm(12, 0, 0.15)
names(subject_shift) <- sprintf("P%02d", 1:12)

effect_window <- dat$condition == "A" & dat$time >= 25 & dat$time <= 38

dat$value <- subject_shift[dat$participant] +
  rnorm(nrow(dat), 0, 0.18) +
  ifelse(effect_window, 1.20, 0)
```

# Prepare data

```{r}
prepared <- prepare_gazepoint_timecourse_test_data(
  data = dat,
  outcome_col = "value",
  time_col = "time",
  condition_col = "condition",
  participant_col = "participant",
  condition_a = "A",
  condition_b = "B"
)

head(prepared)
```

# Run the test

```{r}
result <- run_gazepoint_cluster_permutation(
  data = prepared,
  outcome_col = "value",
  time_col = "time",
  condition_col = "condition",
  participant_col = "participant",
  condition_a = "A",
  condition_b = "B",
  n_permutations = 199,
  cluster_forming_alpha = 0.05,
  cluster_alpha = 0.05,
  seed = 123
)

result
```

# Summarize clusters

```{r}
summarize_gazepoint_time_clusters(result)
```

# Plot result

```{r}
plot_gazepoint_cluster_permutation(result)
```

# Current limitations

This prototype does not yet support:

- between-subject designs
- more than two conditions
- factorial designs
- mixed-effects permutation models
- threshold-free cluster enhancement
- exact onset or offset inference
- trial-level exchangeability schemes
- irregular or incomplete time grids



# Recommended wording

Use cautious wording such as:

> The cluster-based permutation test indicated evidence of a condition difference in the tested time course.

Avoid stronger wording such as:

> The effect began at time X.

# Development status

This article documents an experimental branch prototype. It should not be interpreted as a finalized inferential layer until additional validation, review, and documentation are completed.
