---
title: "Cluster-permutation workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cluster-permutation workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4,
  warning = FALSE,
  message = FALSE
)
```

# Purpose

This article demonstrates the conservative cluster-permutation workflow in `gpbiometrics` for two-condition, one-dimensional time-course data.

The workflow is intentionally narrow: participant-level time courses, two conditions, one temporal dimension, cluster mass based on time-wise test statistics, permutation-based global inference, and descriptive interpretation of cluster timing.

The goal is not to replace specialist EEG, MEG, pupillometry, or mixed-model permutation toolboxes. The goal is to provide a transparent, reproducible, and guarded workflow for Gazepoint-derived time-course signals.

# Interpretation warning

A significant cluster-permutation result indicates evidence against the global null hypothesis of no condition difference anywhere in the tested time range, under the implemented permutation scheme.

It does **not** prove that an effect begins or ends exactly at the reported cluster boundaries. Cluster timing should be described as a descriptive interval, not as a precise physiological, attentional, or cognitive onset or offset.

```{r}
library(gpbiometrics)
```

# Simulate example data

We first create a small synthetic two-condition time-course data set. The example is deliberately synthetic so that the vignette can run without private participant data.

```{r}
set.seed(123)

dat <- simulate_gazepoint_cluster_timecourse_data(
  n_subjects = 12,
  n_time = 60,
  effect_start = 25,
  effect_end = 38,
  seed = 123
)

head(dat)
```

# Audit the time-course grid

Cluster-permutation tests require a consistent participant-by-condition-by-time grid. The audit helper checks whether the expected cells are present and whether duplicate cells exist.

```{r}
grid_audit <- audit_gazepoint_timecourse_grid(
  data = dat,
  subject = "subject",
  condition = "condition",
  time = "time",
  value = "value"
)

grid_audit
```

# Diagnose the design

The design diagnostic is a pre-analysis guardrail. It checks whether the data match the current supported scope: two conditions, repeated participant observations, and a common time grid.

```{r}
design_diagnostic <- diagnose_gazepoint_cluster_design(
  data = dat,
  subject = "subject",
  condition = "condition",
  time = "time",
  value = "value",
  design = "within"
)

design_diagnostic
```

# Prepare participant-level time-course data

The preparation helper keeps the analysis at the participant-by-condition-by-time level. If raw trial-level samples are supplied, repeated cells can be aggregated before the cluster test.

```{r}
prepared <- prepare_gazepoint_timecourse_test_data(
  data = dat,
  outcome_col = "value",
  time_col = "time",
  condition_col = "condition",
  participant_col = "subject"
)

head(prepared)
```

# Run the cluster-permutation test

The example uses a small number of permutations so the vignette builds quickly. For real analyses, use more permutations and report the exact settings.

```{r}
cluster_result <- run_gazepoint_cluster_permutation(
  data = prepared,
  outcome_col = "value",
  time_col = "time",
  condition_col = "condition",
  participant_col = "participant",
  design = "within",
  condition_a = "A",
  condition_b = "B",
  n_permutations = 199,
  cluster_forming_alpha = 0.05,
  cluster_alpha = 0.05,
  seed = 123
)

cluster_result
```

# Summarize clusters

The cluster summary reports detected clusters, cluster mass, and permutation p-values. The time range is useful for description, but it should not be interpreted as an exact onset or offset.

```{r}
cluster_summary <- summarize_gazepoint_time_clusters(cluster_result)

cluster_summary
```

# Plot the result

The plotting helper overlays the two time courses and the detected cluster information. Use the plot to communicate the pattern, not to claim exact temporal localization.

```{r}
plot_gazepoint_cluster_permutation(cluster_result)
```

# Inspect the null distribution

The null-distribution plot shows the permutation distribution of the maximum cluster mass. This helps users understand the family-wise comparison used to evaluate observed clusters.

```{r}
plot_gazepoint_cluster_null_distribution(cluster_result)
```

# Check threshold sensitivity

Cluster-forming thresholds affect which adjacent time points enter clusters. A sensitivity check helps document whether the conclusion is highly dependent on one arbitrary threshold.

```{r}
threshold_sensitivity <- run_gazepoint_cluster_threshold_sensitivity(
  data = prepared,
  dv = "value",
  time = "time",
  condition = "condition",
  subject = "participant",
  thresholds = c(0.025, 0.05, 0.10),
  cluster_alpha = 0.05,
  seed = 123,
  design = "within",
  condition_a = "A",
  condition_b = "B",
  n_permutations = 99
)

threshold_sensitivity
```

# Generate conservative report text

The reporting helper produces cautious wording. It avoids treating cluster boundaries as precise onsets or offsets.

```{r}
cluster_report <- report_gazepoint_cluster_permutation(cluster_result)

cluster_report
```

# Export results

For reproducibility, the result object can be exported to a temporary folder. In a real project, use a project-level output folder such as `outputs/cluster_permutation/`.

```{r}
export_dir <- tempfile("gpbiometrics_cluster_export_")
dir.create(export_dir)

exported_files <- export_gazepoint_cluster_results(
  cluster_result,
  path = export_dir,
  prefix = "cluster_example",
  overwrite = TRUE
)

exported_files
```

# External interoperability

`gpbiometrics` also includes export helpers for users who need to continue in specialist software. These helpers prepare data and metadata; they do not claim to run or validate external models.

```{r}
mne_input <- export_gazepoint_mne_cluster_input(
  data = prepared,
  outcome_col = "value",
  time_col = "time",
  condition_col = "condition",
  participant_col = "participant",
  condition_a = "A",
  condition_b = "B"
)

names(mne_input)
head(mne_input$difference_matrix[, 1:min(6, ncol(mne_input$difference_matrix))])
```

# Advanced guardrails

Some advanced function names are exported deliberately as guardrails. They fail safely instead of pretending to provide validated ANOVA, mixed-model, TFCE, multidimensional, covariate-adjusted, parallel, or onset/offset cluster inference.

```{r}
try(run_gazepoint_tfce())
```

# Recommended wording

Prefer wording such as:

> A two-condition cluster-permutation test indicated evidence of a condition difference in the tested time course. The reported cluster interval is descriptive and should not be interpreted as a precise onset or offset estimate.

Avoid wording such as:

> The effect began exactly at time point X and ended exactly at time point Y.

# Current scope

The current validated workflow is suitable for conservative two-condition time-course analyses. More complex designs should be handled with specialist tools or treated as future methodological work, not as automatic extensions of this workflow.
