---
title: "Design audit workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Design audit workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Scope

This article shows a conservative design-audit workflow for Gazepoint Biometrics projects.

The purpose is to check whether a study design, export folder, event structure, metadata table, and analysis plan are internally consistent before biometric, gaze, pupil, EDA, PPG, HRV, AOI, or multimodal summaries are interpreted.

The workflow focuses on auditability: condition balance, participant and trial coverage, session comparability, export-schema consistency, event coverage, timing coverage, missing design cells, pipeline readiness, release readiness, and reviewer-facing design reports.

Design audits document whether the dataset is structurally suitable for the planned analysis. They do not establish psychological, clinical, emotional, attentional, stress-related, workload-related, or diagnostic states.

# Workflow overview

A typical design-audit workflow is:

1. Create or import trial metadata.
2. Validate required identifiers and design fields.
3. Audit dataset structure.
4. Audit condition balance and missing design cells.
5. Audit session comparability.
6. Audit export schema and export-folder structure.
7. Audit event coverage and timing grids.
8. Audit preprocessing and analysis-pipeline steps.
9. Create design-coverage plots.
10. Create decision logs, audit sections, and reproducibility outputs.

# Example data

For public documentation, use synthetic metadata and synthetic Gazepoint-like streams rather than private exports.

```{r}
library(gpbiometrics)

bio <- simulate_gazepoint_biometrics(
  n_participants = 12,
  n_trials = 16,
  samples_per_trial = 100
)

eye <- simulate_gazepoint_eye_data(
  n_participants = 12,
  n_trials = 16,
  samples_per_trial = 100
)

master <- data.frame(
  participant_id = rep(sprintf("P%02d", 1:12), each = 16),
  trial_id = rep(seq_len(16), times = 12),
  condition = rep(c("A", "B", "C", "D"), length.out = 192),
  session = rep(c("morning", "afternoon"), length.out = 192),
  screen_id = rep(c("screen_simple", "screen_dense"), length.out = 192),
  block = rep(rep(1:4, each = 4), times = 12)
)
```

# Step 1: validate metadata and identifiers

Start by checking that participant, trial, condition, session, and screen identifiers are present and consistently named.

```{r}
validate_gazepoint_metadata(master)

assert_gazepoint_columns(
  master,
  columns = c(
    "participant_id",
    "trial_id",
    "condition",
    "session",
    "screen_id"
  )
)
```

Column standardization should be performed before combining exports and metadata.

```{r}
bio_std <- standardise_gazepoint_biometric_names(bio)
eye_std <- standardize_gazepoint_column_names(eye)

validate_gazepoint_biometrics(bio_std)
validate_gazepoint_format(eye_std)
```

# Step 2: audit dataset structure

Dataset-structure audits document the number of participants, trials, conditions, sessions, and rows available for analysis.

```{r}
dataset_audit <- audit_gazepoint_dataset_structure(
  master,
  participant_col = "participant_id",
  trial_col = "trial_id",
  condition_col = "condition"
)

dataset_audit
```

Export inventories are useful when the design audit starts from a folder of Gazepoint exports.

```{r}
synthetic_dir <- file.path(tempdir(), "gpbiometrics_design_audit_exports")
dir.create(synthetic_dir, recursive = TRUE, showWarnings = FALSE)

write.csv(bio_std, file.path(synthetic_dir, "synthetic_biometrics.csv"), row.names = FALSE)
write.csv(eye_std, file.path(synthetic_dir, "synthetic_eye.csv"), row.names = FALSE)
write.csv(master, file.path(synthetic_dir, "synthetic_master.csv"), row.names = FALSE)

export_profile <- profile_gazepoint_export_folder(synthetic_dir)
export_inventory <- summarize_gazepoint_export_inventory(synthetic_dir)
```

# Step 3: audit condition balance

Condition-balance audits check whether the intended design cells are represented across participants, trials, sessions, and screens.

```{r}
condition_balance <- audit_gazepoint_condition_balance(
  master,
  participant_col = "participant_id",
  condition_col = "condition",
  trial_col = "trial_id"
)

condition_balance
```

For factorial or crossed designs, include all design variables needed for the planned model.

```{r}
experiment_design <- audit_gazepoint_experiment_design(
  master,
  participant_col = "participant_id",
  trial_col = "trial_id",
  design_cols = c("condition", "session", "screen_id", "block")
)

experiment_design
```

# Step 4: audit session comparability

Session comparability checks help identify whether recording sessions, blocks, or screen versions differ in sample counts, trial coverage, or available measurement streams.

```{r}
session_audit <- audit_gazepoint_session_comparability(
  master,
  session_col = "session",
  participant_col = "participant_id",
  trial_col = "trial_id",
  condition_col = "condition"
)

session_audit
```

Session comparability is a structural data-quality question. It should be reported before substantive modelling.

# Step 5: audit export schema


Export-schema audits document whether expected columns and measurement channels are available across files.

```{r}
schema_audit <- audit_gazepoint_export_schema(
  bio_std,
  required_cols = c("participant_id", "trial_id", "TIME_MS")
)

active_channels <- detect_active_biometric_channels(bio_std)
timebase <- detect_gazepoint_biometric_timebase(bio_std)

schema_audit
active_channels
timebase
```

When multiple export folders or batches are present, compare their profiles before pooling them.

```{r}
compare_gazepoint_export_profiles(
  list(
    batch_1 = export_profile,
    batch_2 = export_profile
  )
)
```

# Step 6: audit timing and sampling coverage

Timing audits check whether sample spacing, time resets, and time-course grids are compatible with event-locked or windowed analysis.

```{r}
sampling_audit <- assess_gazepoint_sampling_irregularity(
  bio_std,
  time_col = "TIME_MS",
  participant_col = "participant_id",
  trial_col = "trial_id"
)

time_resets <- audit_gazepoint_time_resets(
  bio_std,
  time_col = "TIME_MS",
  participant_col = "participant_id"
)

timecourse_grid <- audit_gazepoint_timecourse_grid(
  bio_std,
  time_col = "TIME_MS",
  participant_col = "participant_id",
  trial_col = "trial_id"
)
```

# Step 7: audit event coverage

Event coverage checks whether task markers, TTL events, or screen-onset markers are present and sufficiently complete for event-locked analysis.

```{r}
events <- extract_gazepoint_ttl_events(
  bio_std,
  event_col = "EVENT",
  time_col = "TIME_MS",
  participant_col = "participant_id"
)

event_coverage <- audit_gazepoint_event_coverage(
  events,
  participant_col = "participant_id",
  event_col = "EVENT"
)

event_coverage
```

Event coverage should be checked before aligning biometric and gaze streams.

```{r}
event_matches <- match_gazepoint_events_to_biometrics(
  events = events,
  biometrics = bio_std,
  event_time_col = "TIME_MS",
  biometric_time_col = "TIME_MS",
  participant_col = "participant_id",
  tolerance_ms = 100
)
```

# Step 8: audit pipeline readiness

Pipeline audits check whether planned preprocessing and analysis steps have corresponding inputs, outputs, and decision records.

```{r}
pipeline_steps <- data.frame(
  step = c(
    "import",
    "validation",
    "quality_control",
    "event_alignment",
    "feature_extraction",
    "model_preparation",
    "reporting"
  ),
  status = c(
    "complete",
    "complete",
    "planned",
    "planned",
    "planned",
    "planned",
    "planned"
  )
)

pipeline_audit <- audit_gazepoint_pipeline_steps(pipeline_steps)

release_audit <- audit_gazepoint_release_readiness(
  pipeline_steps = pipeline_steps,
  required_steps = c(
    "import",
    "validation",
    "quality_control",
    "feature_extraction",
    "reporting"
  )
)
```

# Step 9: create design-coverage plots

Design-coverage plots are useful for checking imbalance, missing cells, and sparse combinations before modelling.

```{r}
plot_gazepoint_design_coverage(
  master,
  x = "condition",
  fill = "session",
  participant_col = "participant_id"
)
```

Coverage plots should be used as audit displays, not as evidence of substantive effects.

# Step 10: create design-audit report outputs


The audit outputs can be combined into a decision log, QC supplement, audit report section, and reproducibility statement.

```{r}
decision_log <- create_gazepoint_analysis_decision_log(
  decisions = data.frame(
    step = c(
      "metadata_validation",
      "dataset_structure",
      "condition_balance",
      "session_comparability",
      "export_schema",
      "event_coverage",
      "pipeline_readiness"
    ),
    decision = c(
      "Required identifiers and design fields were checked before analysis.",
      "Participant, trial, condition, and session coverage were audited.",
      "Condition balance and missing design cells were reviewed.",
      "Session comparability was audited before pooling sessions.",
      "Export-schema and active-channel checks were completed.",
      "Event coverage and event-to-sample matching were checked before event locking.",
      "Pipeline and release-readiness checks were documented."
    )
  )
)

audit_section <- create_gazepoint_audit_report_section(
  qc_overview = dataset_audit,
  decision_log = decision_log
)

audit_index <- create_gazepoint_audit_index(
  qc_overview = dataset_audit,
  decision_log = decision_log
)

qc_supplement <- create_gazepoint_qc_supplement(
  qc_overview = dataset_audit,
  decision_log = decision_log
)

repro_statement <- create_gazepoint_reproducibility_statement(
  package = "gpbiometrics",
  public_example = "synthetic design-audit example",
  decision_log = decision_log
)
```

# Recommended reporting language

Use precise audit language:

- Report design variables and required identifiers before analysis.
- Report participant, trial, condition, session, block, and screen coverage.
- Report missing design cells and sparse combinations.
- Report session comparability and export-schema consistency.
- Report event coverage, timing-grid checks, and event-matching tolerances.
- Report whether the analysis pipeline had the required inputs, outputs, and decision records.
- Avoid using design-audit outputs as evidence of psychological, clinical, emotional, attentional, stress-related, workload-related, diagnostic, or substantive effects.

# Minimal checklist

Before modelling biometric or multimodal outcomes, confirm that the project has:

- validated participant and trial identifiers;
- validated design variables and metadata fields;
- checked participant, trial, condition, session, block, and screen coverage;
- identified missing or sparse design cells;
- audited export-schema consistency and active measurement channels;
- checked sampling irregularity, time resets, and timing grids;
- audited event coverage before event-locked analysis;
- checked pipeline and release readiness;
- created design-coverage plots;
- documented all audit decisions in a decision log.

# Next steps

After the design audit, prepare toolbox-bridge documentation for exporting gpbiometrics outputs to external analysis ecosystems and then expand the plot gallery.
