---
title: "EDA and SCR visual diagnostics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EDA and SCR visual diagnostics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.6,
  dpi = 96
)
set.seed(20260710)
```

# Purpose

This article demonstrates a plot-rich visual diagnostics workflow for electrodermal activity (EDA), galvanic skin response (GSR), and candidate skin conductance responses (SCRs) in `gpbiometrics` projects.

The workflow is intentionally diagnostic. The figures show signal continuity, unit/range checks, tonic-phasic separation, candidate event marking, response-window summaries, threshold sensitivity, and compact report-style displays. They should not be interpreted as direct evidence of emotion, stress, attention, workload, health status, clinical state, diagnosis, or psychological response.

# Synthetic EDA/GSR stream

The example uses synthetic data so that the article is reproducible and does not depend on private recordings.

```{r synthetic-eda-data}
alpha_scr <- function(time_s, onset_s, amplitude, rise = 1.1, decay = 3.4) {
  x <- pmax(0, time_s - onset_s)
  response <- amplitude * (exp(-x / decay) - exp(-x / rise))
  response[x <= 0] <- 0
  response
}

participants <- sprintf("P%02d", 1:8)
trials <- seq_len(10)
samples_per_trial <- 360
time_s <- seq(0, 24, length.out = samples_per_trial)

trial_grid <- expand.grid(
  participant_id = participants,
  trial_id = trials,
  KEEP.OUT.ATTRS = FALSE,
  stringsAsFactors = FALSE
)

make_trial <- function(participant_id, trial_id) {
  condition <- ifelse(trial_id %% 2 == 0, "dense", "simple")
  participant_shift <- match(participant_id, participants) / 30
  trial_shift <- trial_id / 120
  tonic <- 1.2 + participant_shift + trial_shift + 0.08 * sin(2 * pi * time_s / 24)
  event_onsets <- c(4, 11, 18)
  amplitude_base <- ifelse(condition == "dense", 0.16, 0.11)
  amplitudes <- amplitude_base + rnorm(length(event_onsets), 0, 0.015)
  phasic <- Reduce(`+`, Map(function(onset, amp) alpha_scr(time_s, onset, amp), event_onsets, amplitudes))
  gsr_us <- tonic + phasic + rnorm(length(time_s), 0, 0.015)
  missing <- runif(length(time_s)) < ifelse(condition == "dense", 0.012, 0.007)
  gsr_us[missing] <- NA_real_
  data.frame(
    participant_id = participant_id,
    trial_id = trial_id,
    condition = condition,
    time_s = time_s,
    gsr_us = gsr_us,
    tonic_true = tonic,
    phasic_true = phasic,
    stringsAsFactors = FALSE
  )
}

eda <- do.call(
  rbind,
  Map(make_trial, trial_grid$participant_id, trial_grid$trial_id)
)

head(eda)
```

# GSR signal trace

A single-trial trace is useful for checking continuity, missing stretches, range, and gross artifacts before event extraction.

```{r gsr-trace, fig.cap = "Synthetic GSR trace for one participant-trial segment."}
one_trial <- eda[eda$participant_id == "P01" & eda$trial_id == 2, ]

plot(
  one_trial$time_s,
  one_trial$gsr_us,
  type = "l",
  xlab = "Time (s)",
  ylab = "Conductance (uS)",
  main = "GSR signal trace"
)
```

# Unit and range check

A range plot helps identify unusual scaling or unit-conversion problems. This is a technical plausibility check, not a substantive result.

```{r unit-range-check, fig.cap = "Synthetic GSR conductance range check."}
hist(
  eda$gsr_us,
  breaks = 30,
  xlab = "Conductance (uS)",
  main = "GSR unit and range check"
)
abline(v = median(eda$gsr_us, na.rm = TRUE), lty = 2)
```

# Tonic-phasic decomposition

The example below uses a simple moving-average decomposition for visualization. In formal workflows, the decomposition method and parameters should be documented.

```{r tonic-phasic-decomposition, fig.cap = "Synthetic tonic-phasic decomposition display."}
smooth_tonic <- stats::filter(one_trial$gsr_us, rep(1 / 51, 51), sides = 2)
smooth_tonic <- as.numeric(smooth_tonic)
smooth_tonic[is.na(smooth_tonic)] <- approx(
  x = which(!is.na(smooth_tonic)),
  y = smooth_tonic[!is.na(smooth_tonic)],
  xout = which(is.na(smooth_tonic)),
  rule = 2
)$y

phasic_est <- one_trial$gsr_us - smooth_tonic

op <- par(mfrow = c(2, 1), mar = c(3.5, 4, 2.5, 1))
plot(one_trial$time_s, one_trial$gsr_us, type = "l", xlab = "Time (s)", ylab = "GSR (uS)", main = "Observed GSR and tonic estimate")
lines(one_trial$time_s, smooth_tonic, lty = 2)
plot(one_trial$time_s, phasic_est, type = "l", xlab = "Time (s)", ylab = "Phasic estimate", main = "Estimated phasic component")
abline(h = 0, lty = 2)
par(op)
```

# Candidate SCR event markers

Candidate event markers document the detection rule and provide a visual audit trail for later review.

```{r scr-event-markers, fig.cap = "Synthetic candidate SCR event markers."}
event_onsets <- c(4, 11, 18)
candidate_events <- do.call(
  rbind,
  lapply(event_onsets, function(onset) {
    idx <- which(one_trial$time_s >= onset & one_trial$time_s <= onset + 5)
    peak_idx <- idx[which.max(phasic_est[idx])]
    data.frame(
      onset_s = onset,
      peak_s = one_trial$time_s[peak_idx],
      amplitude = phasic_est[peak_idx],
      stringsAsFactors = FALSE
    )
  })
)
candidate_events$latency_s <- candidate_events$peak_s - candidate_events$onset_s

plot(
  one_trial$time_s,
  phasic_est,
  type = "l",
  xlab = "Time (s)",
  ylab = "Phasic estimate",
  main = "Candidate SCR events"
)
abline(h = 0, lty = 2)
abline(v = candidate_events$onset_s, lty = 3)
points(candidate_events$peak_s, candidate_events$amplitude, pch = 16)
text(candidate_events$peak_s, candidate_events$amplitude, labels = seq_len(nrow(candidate_events)), pos = 3)
candidate_events
```

# Latency and recovery summaries

Latency and recovery summaries are useful for checking event-window extraction. They should be reported as algorithmic summaries unless the study design supports stronger inference.

```{r latency-recovery-summary, fig.cap = "Synthetic SCR latency and recovery summaries."}
event_summary <- do.call(
  rbind,
  lapply(seq_len(nrow(trial_grid)), function(i) {
    condition <- ifelse(trial_grid$trial_id[i] %% 2 == 0, "dense", "simple")
    amplitude_base <- ifelse(condition == "dense", 0.16, 0.11)
    data.frame(
      participant_id = trial_grid$participant_id[i],
      trial_id = trial_grid$trial_id[i],
      condition = condition,
      event = seq_along(event_onsets),
      onset_s = event_onsets,
      amplitude = pmax(0.02, amplitude_base + rnorm(length(event_onsets), 0, 0.025)),
      latency_s = pmax(0.4, 1.6 + rnorm(length(event_onsets), 0, 0.25)),
      recovery_s = pmax(1.0, 4.2 + rnorm(length(event_onsets), 0, 0.55)),
      stringsAsFactors = FALSE
    )
  })
)

op <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
boxplot(latency_s ~ condition, data = event_summary, xlab = "Synthetic condition", ylab = "Latency (s)", main = "SCR latency")
boxplot(recovery_s ~ condition, data = event_summary, xlab = "Synthetic condition", ylab = "Recovery time (s)", main = "SCR recovery")
par(op)
```

# Threshold sensitivity

Threshold sensitivity plots show how candidate-response counts change when the amplitude threshold changes.

```{r threshold-sensitivity, fig.cap = "Synthetic SCR threshold sensitivity display."}
thresholds <- seq(0.04, 0.20, by = 0.02)
sens <- do.call(
  rbind,
  lapply(thresholds, function(th) {
    aggregate(
      event_summary$amplitude >= th,
      by = list(condition = event_summary$condition),
      FUN = mean
    ) |>
      transform(threshold = th)
  })
)
names(sens)[2] <- "detection_rate"

simple <- sens[sens$condition == "simple", ]
dense <- sens[sens$condition == "dense", ]

plot(
  simple$threshold,
  simple$detection_rate * 100,
  type = "b",
  ylim = c(0, 100),
  xlab = "Amplitude threshold",
  ylab = "Candidate events retained (%)",
  main = "SCR threshold sensitivity"
)
lines(dense$threshold, dense$detection_rate * 100, type = "b", pch = 17, lty = 2)
legend("topright", legend = c("simple", "dense"), lty = c(1, 2), pch = c(1, 17), bty = "n")
```

# Compact EDA/SCR visual diagnostics dashboard

A compact dashboard can combine range, signal, event, and threshold diagnostics for review or supplementary reporting.

```{r compact-eda-dashboard, fig.cap = "Compact synthetic EDA/SCR visual diagnostics dashboard."}
op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

hist(eda$gsr_us, breaks = 25, xlab = "Conductance (uS)", main = "Range check")

plot(one_trial$time_s, one_trial$gsr_us, type = "l", xlab = "Time (s)", ylab = "GSR (uS)", main = "Trace")

plot(one_trial$time_s, phasic_est, type = "l", xlab = "Time (s)", ylab = "Phasic", main = "Candidate SCRs")
points(candidate_events$peak_s, candidate_events$amplitude, pch = 16)

plot(simple$threshold, simple$detection_rate * 100, type = "b", ylim = c(0, 100), xlab = "Threshold", ylab = "Retained (%)", main = "Sensitivity")
lines(dense$threshold, dense$detection_rate * 100, type = "b", pch = 17, lty = 2)

par(op)
```

# Relation to gpbiometrics helpers

The same reporting pattern can be paired with package-level helpers when project data use the expected column contracts.

```{r helper-map, echo = FALSE}
helper_map <- data.frame(
  task = c(
    "GSR unit audit",
    "EDA decomposition",
    "SCR event detection",
    "Latency and recovery summaries",
    "Threshold sensitivity",
    "Report outputs"
  ),
  representative_helpers = c(
    "audit_gazepoint_gsr_units(); convert_gazepoint_gsr_to_conductance()",
    "decompose_gazepoint_eda(); plot_gazepoint_eda_decomposition()",
    "detect_gazepoint_scr_events(); detect_gazepoint_scr_peaks(); plot_gazepoint_scr_events()",
    "compute_gazepoint_scr_latency(); extract_gazepoint_scr_recovery_times()",
    "run_gazepoint_scr_threshold_sensitivity(); plot_gazepoint_scr_specification_curve()",
    "create_gazepoint_qc_supplement(); create_gazepoint_biometrics_report_tables()"
  ),
  stringsAsFactors = FALSE
)

knitr::kable(helper_map)
```

```{r package-helper-pattern, eval = FALSE}
library(gpbiometrics)

gsr_units <- audit_gazepoint_gsr_units(
  biometric_data,
  gsr_col = "GSR"
)

eda_conductance <- convert_gazepoint_gsr_to_conductance(
  biometric_data,
  gsr_col = "GSR",
  output_col = "GSR_US"
)

eda_decomp <- decompose_gazepoint_eda(
  eda_conductance,
  eda_col = "GSR_US",
  time_col = "TIME_MS",
  participant_col = "participant_id",
  trial_col = "trial_id"
)

scr_events <- detect_gazepoint_scr_events(
  eda_decomp,
  phasic_col = "eda_phasic",
  time_col = "TIME_MS",
  participant_col = "participant_id",
  trial_col = "trial_id"
)

plot_gazepoint_eda_decomposition(eda_decomp)
plot_gazepoint_scr_events(eda_decomp, events = scr_events)
```

# Reporting recommendation

For manuscripts, technical supplements, and reviewer-facing reports, describe these figures as EDA/GSR/SCR diagnostics, signal-processing checks, event-detection displays, threshold-sensitivity checks, and reproducibility outputs. Avoid stronger psychological, clinical, affective, workload-related, or stress-related interpretation unless supported by validation, controls, and modelling evidence.
