---
title: "Using gpbiometrics with eyetrackingR, PupillometryR, and gazeR"
description: "Compare and prepare Gazepoint data for three optional R eye-tracking and pupillometry ecosystems."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using gpbiometrics with eyetrackingR, PupillometryR, and gazeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

if (file.exists("DESCRIPTION") && dir.exists("R")) {
  devtools::load_all(".", quiet = TRUE)
} else {
  library(gpbiometrics)
}
```

# Purpose

`gpbiometrics` provides optional bridges to `eyetrackingR`,
`PupillometryR`, and gazeR. These bridges standardize source columns, preserve
an audit record, and optionally construct an external-package object when that
package is installed.

The bridge is a data-preparation boundary. Downstream modelling,
visualization, and package-specific assumptions remain the responsibility of
the selected external package.

# Which bridge should be used?

```{r bridge-table}
bridge_functions <- data.frame(
  target = c(
    "eyetrackingR",
    "PupillometryR",
    "gazeR"
  ),
  helper = c(
    "prepare_gazepoint_eyetrackingr_input",
    "prepare_gazepoint_pupillometryr_input",
    "prepare_gazepoint_gazer_input"
  ),
  primary_focus = c(
    "AOI and time-window analysis",
    "Pupil preprocessing and time-course analysis",
    "Gaze, pupil, blink, and AOI preparation"
  ),
  stringsAsFactors = FALSE
)

bridge_functions$available <-
  bridge_functions$helper %in%
  getNamespaceExports("gpbiometrics")

bridge_functions

stopifnot(all(bridge_functions$available))
```

# Common source contract

A bridge-ready table should preserve:

- participant identifier;
- trial or media identifier;
- ordered time variable with a documented unit;
- gaze coordinates where required;
- pupil values where required;
- validity or track-loss information;
- blink information where available;
- AOI membership or the columns required to construct AOIs;
- condition and stimulus metadata.

Do not remove original Gazepoint columns until the conversion audit has been
reviewed.

# eyetrackingR preparation

```{r eyetrackingr-signature}
args(prepare_gazepoint_eyetrackingr_input)
```

```{r eyetrackingr-example, eval=FALSE}
eyetrackingr_input <- prepare_gazepoint_eyetrackingr_input(
  data = gaze_samples,
  ...
)
```

Review participant, trial, time, track-loss, and logical AOI columns before
constructing an `eyetrackingR_data` object.

# PupillometryR preparation

```{r pupillometryr-signature}
args(prepare_gazepoint_pupillometryr_input)
```

```{r pupillometryr-example, eval=FALSE}
pupillometryr_input <- prepare_gazepoint_pupillometryr_input(
  data = pupil_samples,
  ...
)
```

Document whether binocular or monocular pupil values were selected, how
invalid values were masked, whether time was converted to milliseconds, and
whether blink periods were retained or removed.

# gazeR preparation

```{r gazer-signature}
args(prepare_gazepoint_gazer_input)
```

```{r gazer-example, eval=FALSE}
gazer_input <- prepare_gazepoint_gazer_input(
  data = gaze_samples,
  ...
)
```

gazeR preparation may include monocular or binocular coordinates, pupil
columns, validity values, blink flags, time conversion, and
irregular-sampling checks.

# Cross-bridge comparison

| Decision | Question |
|---|---|
| Analysis target | AOI proportions, pupil time course, blink processing, or gaze processing? |
| Time unit | Does the target package require milliseconds? |
| Validity rule | How will invalid coordinates or pupil values be represented? |
| Trial structure | Are participant and trial identifiers complete and stable? |
| AOI representation | Logical AOI columns or one categorical AOI field? |
| External dependency | Is object construction required or only a data frame? |
| Auditability | Are renamed, masked, omitted, and retained columns recorded? |

# Optional object construction

```{r object-construction, eval=FALSE}
prepared <- prepare_gazepoint_eyetrackingr_input(
  data = gaze_samples,
  ...
)

# Inspect the prepared data, mapping, audit, and settings before
# constructing an external-package object.
```

# Reporting checklist

Report:

- source Gazepoint export and package versions;
- selected bridge and target-package version;
- participant, trial, time, gaze, pupil, AOI, validity, and blink mappings;
- time-unit conversion;
- invalid-sample masking;
- binocular or monocular processing;
- dropped rows or columns;
- irregular-sampling decisions;
- whether an external-package object was constructed.

# Interpretation guardrails

Conversion does not validate downstream model assumptions. Pupil, gaze, AOI,
blink, and track-loss outputs should be interpreted within the experimental
design and documented preprocessing workflow.
