---
title: "Testing interoperability across external package versions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing interoperability across external package versions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)

library(gpbiometrics)
```

## Purpose

`gpbiometrics` prepares data and metadata for several external R and Python
ecosystems. These integrations can be affected when external packages change
their accepted columns, object structures, runtime requirements, or supported
versions.

The interoperability audit checks the contracts controlled by
`gpbiometrics`. It does not attempt to prove that external analytical
algorithms produce identical numerical results across releases.

The maintained checks focus on whether:

- declared bridge functions remain exported;
- optional dependencies are detected safely;
- installed versions can be recorded and compared with tested floors;
- Python runtime availability is reported explicitly;
- external-package absence does not prevent the core package from loading;
- compatibility findings can be written without source or participant data.

## Interoperability manifest

The manifest provides a machine-readable description of the maintained
integration surface.

```{r manifest}
manifest <- gazepoint_interoperability_manifest()

manifest[
  ,
  c(
    "target",
    "ecosystem",
    "dependency",
    "minimum_tested_version",
    "version_policy",
    "test_group",
    "optional"
  )
]
```

The manifest currently covers:

- `eyetrackingR`;
- `PupillometryR`;
- `gazeR`;
- MNE-Python;
- `pylsl`;
- BioSPPy;
- HeartPy;
- pyHRV;
- BIDS;
- NumPy;
- pandas.

A minimum tested version identifies the oldest version exercised by the
maintained compatibility workflow. It does not establish that every earlier
release is incompatible.

For R packages without a historical floor, the policy is
`"current-installed"`. Their installed versions are recorded, but no claim is
made about an untested lower boundary.

## Dependency-free audit

A dependency-free audit checks all bridge exports and records R-package
availability without initializing Python.

```{r dependency-free-audit}
audit <- audit_gazepoint_interoperability_versions(
  include_python = FALSE,
  strict = TRUE
)

print(audit)
```

```{r audit-summary}
audit$summary
```

A compact target-level view is usually sufficient for review.

```{r audit-results}
audit$results[
  ,
  c(
    "target",
    "dependency",
    "installed_version",
    "minimum_tested_version",
    "status",
    "pass",
    "needs_review"
  )
]
```

Typical statuses include:

- `available`: the installed dependency meets its declared tested floor;
- `available_unpinned`: the dependency is installed, but no historical floor
  has been declared;
- `declared`: a standard-level contract, such as BIDS, is registered;
- `not_checked`: Python inspection was deliberately disabled;
- `runtime_unavailable`: Python could not be initialized;
- `missing_dependency`: an optional dependency is absent;
- `below_minimum`: an installed dependency is below the declared tested floor;
- `missing_bridge`: a declared `gpbiometrics` bridge export is unavailable.

Missing optional dependencies and deliberately disabled Python checks are
review conditions rather than core-package failures. Missing bridge exports
and installed versions below declared floors fail the strict audit.

## Python-enabled audit

A Python-enabled audit uses the active `reticulate` environment.

```{r python-audit, eval=FALSE}
Sys.setenv(
  RETICULATE_PYTHON =
    "C:/path/to/python.exe"
)

python_audit <-
  audit_gazepoint_interoperability_versions(
    include_python = TRUE,
    strict = TRUE
  )

python_audit$summary

python_audit$results[
  python_audit$results$ecosystem == "Python",
  c(
    "target",
    "installed_version",
    "minimum_tested_version",
    "runtime_version",
    "status",
    "pass"
  )
]
```

Python distribution versions are queried through `importlib.metadata`.
The audit does not import every analytical package merely to obtain its
version.

Actual bridge tests remain responsible for checking object construction,
metadata preservation, input structure, time units, and safe failure
behaviour.

## Writing a compatibility report

Compatibility reports contain aggregate environment and target information
only.

```{r write-report}
output_dir <- tempfile(
  "gpbiometrics-interoperability-"
)

files <- write_gazepoint_interoperability_audit(
  audit,
  output_dir = output_dir
)

basename(files)
```

The generated files contain:

- target-level results;
- an aggregate pass/review summary;
- R, Python, package, platform, and operating-system metadata;
- the manifest used by the audit.

They do not contain imported signals, participant identifiers, source
filenames, or input-data paths.

```{r cleanup-report, include=FALSE}
unlink(
  output_dir,
  recursive = TRUE,
  force = TRUE
)
```

## Command-line audit

From the package repository, the maintained command-line runner can be used
without writing results into the repository.

```{r command-line-runner, eval=FALSE}
system2(
  "Rscript",
  c(
    "tools/run-interoperability-version-audit.R",
    shQuote(
      "C:/external-output/gpbiometrics-interoperability"
    )
  )
)
```

The same location can be supplied through an environment variable.

```{r command-line-environment, eval=FALSE}
Sys.setenv(
  GPBIOMETRICS_INTEROP_OUTPUT =
    "C:/external-output/gpbiometrics-interoperability",
  GPBIOMETRICS_INTEROP_PYTHON =
    "true"
)
```

## Continuous-integration matrix

The dedicated GitHub Actions workflow runs three R bridge jobs and two Python
profiles.

The R jobs exercise the eye-tracking bridge contracts on:

- Windows;
- Ubuntu;
- macOS.

The Python jobs exercise:

- a declared floor environment;
- a current-release environment.

The workflow stores compatibility CSV files as temporary GitHub Actions
artifacts. These artifacts provide an auditable runtime record without adding
generated compatibility files to the repository.

## Interpretation boundary

The audit supports statements such as:

- a declared input contract remains available;
- an optional dependency is installed or absent;
- a tested version floor is met;
- expected metadata and bridge exports remain present;
- a runtime environment requires review.

It does not support claims that:

- every version of an external package is compatible;
- external numerical implementations are equivalent across releases;
- successful object construction validates a scientific analysis;
- an absent optional dependency is a failure of the core package.

Interoperability findings should therefore be interpreted as software-contract
evidence rather than scientific equivalence evidence.
