---
title: "MNE, EEG, and LSL interoperability workflow"
description: "Prepare Gazepoint events and signal channels for MNE, align Gazepoint and EEG clocks, and document LSL/XDF synchronization."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MNE, EEG, and LSL interoperability workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

if (file.exists("DESCRIPTION") && dir.exists("R")) {
  devtools::load_all(".", quiet = TRUE)
} else {
  library(gpbiometrics)
}
```

# Purpose

This article documents an auditable workflow for preparing Gazepoint events and
signal channels for MNE-Python, aligning Gazepoint timestamps with an EEG
clock, and handling LSL or XDF synchronization information.

The interoperability helpers prepare data and synchronization records. They do
not replace EEG preprocessing, event-quality review, or specialist MNE
analysis.

# Workflow map

```{r function-inventory}
workflow_functions <- data.frame(
  stage = c(
    "Event conversion",
    "Channel preparation",
    "EEG clock alignment",
    "LSL stream synchronization",
    "LSL clock-offset estimation",
    "Optional FIF writing",
    "Reproducibility metadata"
  ),
  function_name = c(
    "prepare_gazepoint_mne_events",
    "prepare_gazepoint_mne_input",
    "align_gazepoint_to_eeg",
    "sync_gazepoint_signals_via_lsl",
    "estimate_gazepoint_lsl_clock_offsets",
    "write_gazepoint_mne_fif",
    "session_info_gazepoint"
  ),
  stringsAsFactors = FALSE
)

workflow_functions$available <-
  workflow_functions$function_name %in%
  getNamespaceExports("gpbiometrics")

workflow_functions

stopifnot(all(workflow_functions$available))
```

# Prepare event information

A useful event table should retain:

- participant and recording identifiers;
- event labels or numeric event codes;
- original Gazepoint timestamps;
- the time unit and recording origin;
- TTL validity or marker-channel information;
- a stable event identifier.

```{r mne-event-signature}
args(prepare_gazepoint_mne_events)
```

```{r mne-event-example, eval=FALSE}
mne_events <- prepare_gazepoint_mne_events(
  ...
)
```

Confirm that event order, sample indices, event codes, and timestamp conversion
agree with the experimental log.

# Prepare Gazepoint channels for MNE

```{r mne-input-signature}
args(prepare_gazepoint_mne_input)
```

```{r mne-input-example, eval=FALSE}
mne_input <- prepare_gazepoint_mne_input(
  ...
)
```

Document every exported channel's source column, signal type, physical unit,
sampling rate, missing-data handling, filtering history, and validity rule.

Do not label an HRV validity field as an HRV metric. EDA, pupil, pulse, gaze,
HR, and IBI channels should not be reinterpreted as direct psychological
states.

# Align Gazepoint and EEG clocks

```{r eeg-alignment-signature}
args(align_gazepoint_to_eeg)
```

```{r eeg-alignment-example, eval=FALSE}
alignment <- align_gazepoint_to_eeg(
  ...
)
```

Retain original and adjusted timestamps, estimated offset, estimated drift,
matched event pairs, unmatched events, and residual timing errors.

Alignment should not silently remove events or conceal poor correspondence
between recording clocks.

# Synchronize LSL or XDF streams

```{r lsl-signatures}
args(sync_gazepoint_signals_via_lsl)
args(estimate_gazepoint_lsl_clock_offsets)
```

```{r lsl-example, eval=FALSE}
clock_offsets <- estimate_gazepoint_lsl_clock_offsets(
  ...
)

lsl_sync <- sync_gazepoint_signals_via_lsl(
  ...
)
```

Live clock-offset estimation requires a local `pylsl` installation. Imported
XDF workflows should retain stream names, source identifiers, nominal rates,
original timestamps, corrected timestamps, and any dejittering decision.

# Optional native FIF writing

```{r fif-signature}
args(write_gazepoint_mne_fif)
```

```{r fif-example, eval=FALSE}
fif_result <- write_gazepoint_mne_fif(
  ...
)
```

Native FIF writing requires a local Python and MNE installation. Use a dry run
or temporary output directory first and inspect the resulting channel
metadata.

# Reporting checklist

Report at minimum:

- Gazepoint and EEG acquisition rates;
- event sources and marker definitions;
- timestamp units and clock origins;
- event-matching method;
- estimated offset and drift;
- synchronization residuals;
- unmatched or excluded events;
- resampling, interpolation, or dejittering decisions;
- MNE, Python, pylsl, R, and gpbiometrics versions.

```{r session-info-signature}
args(session_info_gazepoint)
```

# Interpretation guardrails

Synchronization establishes temporal correspondence, not causal or
psychological interpretation. Aligned signals remain measurements whose
interpretation depends on study design, preprocessing, operationalization, and
convergent evidence.
