---
title: "Plot gallery"
description: "Rendered examples of all exported gpbiometrics plot helpers."
output: rmarkdown::html_vignette
---

<!--
%\VignetteIndexEntry{Plot gallery}
%\VignetteEngine{knitr::rmarkdown}
%\VignetteEncoding{UTF-8}
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  results = "hide",
  fig.width = 8,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)

library(gpbiometrics)

.show_plot_result <- function(res) {
  if (inherits(res, "ggplot")) {
    print(res)
    return(invisible(res))
  }

  if (inherits(res, "recordedplot")) {
    replayPlot(res)
    return(invisible(res))
  }

  if (inherits(res, c("grob", "gtable"))) {
    grid::grid.newpage()
    grid::grid.draw(res)
    return(invisible(res))
  }

  if (is.list(res)) {
    plot_like <- Filter(
      function(z) inherits(z, c("ggplot", "recordedplot", "grob", "gtable")),
      res
    )

    if (length(plot_like)) {
      for (z in plot_like) {
        .show_plot_result(z)
      }
    }
  }

  invisible(res)
}

.gp <- function(name) {
  f <- get(name, envir = asNamespace("gpbiometrics"), inherits = FALSE)

  function(...) {
    res <- f(...)
    .show_plot_result(res)
    invisible(res)
  }
}
set.seed(20260630)
```

# Plot gallery

This article showcases all exported `plot_*` helpers in `gpbiometrics` using synthetic Gazepoint-like data.

The example code is hidden so that the page functions as a compact visual gallery. Function names are shown above each rendered example.

The examples are for documentation and visual inspection only. The synthetic signals are not real physiology and should not be interpreted as emotion, stress, cognition, preference, health status, or diagnosis.

```{r synthetic-data}
fs <- 60
time_s <- seq(0, 120, by = 1 / fs)
n <- length(time_s)

condition <- rep(c("simple_clear", "dense_ambiguous"), length.out = n)

ppg <- sin(2 * pi * 1.15 * time_s) +
  0.25 * sin(2 * pi * 0.22 * time_s) +
  0.04 * stats::rnorm(n)

gsr_us <- 2.0 +
  0.05 * sin(2 * pi * 0.015 * time_s) +
  0.02 * stats::rnorm(n)

scr_onsets <- seq(8, 110, by = 14)

for (on in scr_onsets) {
  gsr_us <- gsr_us + ifelse(time_s >= on, 0.12 * exp(-(time_s - on) / 2.5), 0)
}

hr <- 72 + 4 * sin(2 * pi * 0.01 * time_s) + stats::rnorm(n, 0, 0.5)
ibi_ms <- 60000 / hr

demo_signal <- data.frame(
  participant = rep(sprintf("P%02d", 1:12), length.out = n),
  trial = rep(1:12, length.out = n),
  condition = condition,
  time_s = time_s,
  timestamp = time_s,
  MSTIMER = time_s * 1000,
  GSR_US = gsr_us,
  GSR = gsr_us,
  EDA = gsr_us,
  PPG = ppg,
  HR = hr,
  IBI = ibi_ms,
  pupil_left = 3.1 + 0.1 * sin(2 * pi * 0.04 * time_s) + stats::rnorm(n, 0, 0.02),
  pupil_right = 3.1 + 0.1 * sin(2 * pi * 0.04 * time_s) + stats::rnorm(n, 0, 0.02),
  gaze_x = 0.5 + 0.25 * sin(2 * pi * 0.03 * time_s) + stats::rnorm(n, 0, 0.03),
  gaze_y = 0.5 + 0.20 * cos(2 * pi * 0.02 * time_s) + stats::rnorm(n, 0, 0.03),
  TTL = as.integer(abs(time_s %% 10) < (1 / fs)),
  stringsAsFactors = FALSE
)

demo_signal$AOI <- cut(
  demo_signal$gaze_x,
  breaks = c(-Inf, 0.35, 0.50, 0.65, 0.80, Inf),
  labels = c("Instructions", "Claim", "Evidence", "CTA", "Other")
)

demo_signal$aoi <- as.character(demo_signal$AOI)

rr_ms <- 820 +
  45 * sin(seq(0, 8 * pi, length.out = 180)) +
  stats::rnorm(180, 0, 18)

rr_ms <- as.numeric(rr_ms)
rr_ms <- rr_ms[is.finite(rr_ms) & rr_ms > 300 & rr_ms < 2000]

nni_ms <- rr_ms
nni_time_s <- seq_along(nni_ms)

aoi_bio <- expand.grid(
  aoi_label = c("Instructions", "Claim", "Evidence", "CTA"),
  signal = c("GSR", "HR", "pupil_left"),
  condition = c("simple_clear", "dense_ambiguous"),
  stringsAsFactors = FALSE
)

aoi_bio$mean_value <- with(
  aoi_bio,
  ifelse(signal == "GSR", 2.0,
    ifelse(signal == "HR", 72, 3.2)
  ) +
    ifelse(condition == "dense_ambiguous", 0.15, 0) +
    rep(c(0.00, 0.06, 0.10, 0.04), length.out = nrow(aoi_bio))
)

ppg_detection <- .gp("detect_gazepoint_ppg_peaks")(
  data = demo_signal,
  time_col = "time_s",
  signal_col = "PPG"
)

segment_measures <- data.frame(
  segment = seq_len(12),
  segment_id = seq_len(12),
  segment_start_s = seq(0, by = 20, length.out = 12),
  segment_end_s = seq(20, by = 20, length.out = 12),
  bpm = 72 + 4 * sin(seq(0, 2 * pi, length.out = 12)) + stats::rnorm(12, 0, 0.8),
  ibi_ms = 60000 / (72 + 4 * sin(seq(0, 2 * pi, length.out = 12))),
  sdnn = 40 + 5 * sin(seq(0, 2 * pi, length.out = 12)),
  rmssd = 32 + 4 * cos(seq(0, 2 * pi, length.out = 12)),
  quality = runif(12, 0.85, 0.99)
)

segmentwise <- list(
  measures = segment_measures,
  segment_measures = segment_measures,
  summary = segment_measures,
  settings = list(measure = "bpm")
)

class(segmentwise) <- c("gazepoint_ppg_segmentwise", "list")

measures <- data.frame(
  sdnn = 52,
  rmssd = 38,
  sdsd = 36,
  pnn50 = 18,
  lf_norm = 54,
  hf_norm = 46,
  sd1 = 27,
  sd2 = 68
)

scr_peaks <- data.frame(
  onset_time_s = scr_onsets,
  peak_time_s = scr_onsets + 2.0,
  recovery_time_s = scr_onsets + 5.5,
  amplitude = 0.10 + 0.04 * sin(seq_along(scr_onsets)),
  peak_amplitude = 0.10 + 0.04 * sin(seq_along(scr_onsets)),
  condition = rep(c("simple_clear", "dense_ambiguous"), length.out = length(scr_onsets)),
  stringsAsFactors = FALSE
)

spec_curve <- data.frame(
  specification_id = sprintf("spec_%02d", 1:24),
  estimate = seq(-0.08, 0.18, length.out = 24) + stats::rnorm(24, 0, 0.01),
  se = runif(24, 0.015, 0.035),
  p_value = seq(0.20, 0.001, length.out = 24),
  threshold = rep(c("0.01", "0.03", "0.05", "0.07"), length.out = 24),
  baseline = rep(c("none", "pre_event", "trial"), length.out = 24),
  smoothing = rep(c("raw", "ema"), length.out = 24),
  stringsAsFactors = FALSE
)

saccades <- data.frame(
  duration_ms = seq(20, 80, length.out = 80),
  amplitude_deg = seq(0.5, 8, length.out = 80),
  peak_velocity_deg_s = 60 + 45 * log1p(seq(0.5, 8, length.out = 80)) + stats::rnorm(80, 0, 8),
  participant = rep(sprintf("P%02d", 1:10), length.out = 80),
  condition = rep(c("simple_clear", "dense_ambiguous"), length.out = 80),
  stringsAsFactors = FALSE
)
```

# AOI and multimodal plots

## `plot_gazepoint_aoi_biometrics()`

```{r plot-aoi-biometrics}
.gp("plot_gazepoint_aoi_biometrics")(
  x = aoi_bio,
  value_col = "mean_value",
  aoi_col = "aoi_label",
  signal_col = "signal",
  group_col = "condition"
)
```

## `plot_gazepoint_multimodal_timeline()`

```{r plot-multimodal-timeline}
.gp("plot_gazepoint_multimodal_timeline")(
  data = demo_signal,
  time_col = "time_s",
  signal_cols = c("GSR", "PPG", "HR", "IBI", "pupil_left"),
  event_col = "TTL",
  participant_col = "participant",
  trial_col = "trial"
)
```

# Biometric quality and signal plots

## `plot_gazepoint_biometric_quality()`

```{r plot-biometric-quality}
.gp("plot_gazepoint_biometric_quality")(
  data = demo_signal,
  time_col = "time_s",
  signal_cols = c("GSR", "PPG", "HR", "IBI", "pupil_left"),
  group_col = "condition"
)
```

## `plot_gazepoint_biometric_report_dashboard()`

```{r plot-report-dashboard}
.gp("plot_gazepoint_biometric_report_dashboard")(
  data = demo_signal,
  time_col = "time_s",
  signal_cols = c("GSR", "PPG", "HR", "IBI", "pupil_left")
)
```

## `plot_gazepoint_biometric_signals()`

```{r plot-biometric-signals}
.gp("plot_gazepoint_biometric_signals")(
  data = demo_signal,
  time_col = "time_s",
  signal_cols = c("GSR", "PPG", "HR", "IBI", "pupil_left"),
  group_col = "condition"
)
```

## `plot_gazepoint_signal_activity()`

```{r plot-signal-activity}
.gp("plot_gazepoint_signal_activity")(
  data = demo_signal,
  signal_cols = c("GSR", "PPG", "HR", "IBI", "pupil_left")
)
```

## `plot_gazepoint_time_resets()`

```{r plot-time-resets}
.gp("plot_gazepoint_time_resets")(
  data = demo_signal,
  time_col = "time_s"
)
```

# EDA and SCR plots

## `plot_gazepoint_eda_decomposition()`

```{r plot-eda-decomposition}
.gp("plot_gazepoint_eda_decomposition")(
  data = demo_signal,
  time_col = "time_s",
  signal_cols = c("GSR", "PPG", "HR", "IBI", "pupil_left")
)
```

## `plot_gazepoint_eda_gram()`

```{r plot-eda-gram}
.gp("plot_gazepoint_eda_gram")(
  dat = demo_signal,
  time_col = "time_s",
  eda_col = "GSR_US"
)
```

## `plot_gazepoint_scr_events()`

```{r plot-scr-events}
.gp("plot_gazepoint_scr_events")(
  data = demo_signal,
  scr_peaks = scr_peaks,
  time_col = "time_s",
  signal_col = "GSR_US"
)
```

## `plot_gazepoint_scr_specification_curve()`

```{r plot-scr-specification-curve}
.gp("plot_gazepoint_scr_specification_curve")(
  x = spec_curve,
  estimate_col = "estimate",
  specification_col = "specification_id"
)
```

# PPG, HRV, and respiration plots

## `plot_gazepoint_ppg_breathing()`

```{r plot-ppg-breathing}
.gp("plot_gazepoint_ppg_breathing")(
  rr_ms = rr_ms
)
```

## `plot_gazepoint_ppg_peak_detection()`

```{r plot-ppg-peak-detection}
.gp("plot_gazepoint_ppg_peak_detection")(
  detection = ppg_detection
)
```

## `plot_gazepoint_ppg_poincare()`

```{r plot-ppg-poincare}
.gp("plot_gazepoint_ppg_poincare")(
  rr_ms = rr_ms
)
```

## `plot_gazepoint_ppg_segmentwise()`

```{r plot-ppg-segmentwise}
.gp("plot_gazepoint_ppg_segmentwise")(
  segmentwise = segmentwise,
  measure = "bpm"
)
```

## `plot_gazepoint_pyhrv_hr_heatplot()`

```{r plot-pyhrv-hr-heatplot}
.gp("plot_gazepoint_pyhrv_hr_heatplot")(
  nni_ms = nni_ms,
  time_bins = 20,
  hr_bins = 20
)
```

## `plot_gazepoint_pyhrv_radar_chart()`

```{r plot-pyhrv-radar-chart}
.gp("plot_gazepoint_pyhrv_radar_chart")(
  measures = measures
)
```

## `plot_gazepoint_pyhrv_tachogram()`

```{r plot-pyhrv-tachogram}
.gp("plot_gazepoint_pyhrv_tachogram")(
  nni_ms = nni_ms,
  time_s = nni_time_s
)
```

# Eye-movement plot

## `plot_gazepoint_saccade_main_sequence()`

```{r plot-saccade-main-sequence}
.gp("plot_gazepoint_saccade_main_sequence")(
  dat = saccades,
  amplitude_col = "amplitude_deg",
  peak_velocity_col = "peak_velocity_deg_s",
  group_col = "condition"
)
```
