---
title: "PPG and HRV visual diagnostics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PPG and HRV visual diagnostics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.6,
  dpi = 96
)
set.seed(20260710)
```

# Purpose

This article demonstrates a plot-rich visual diagnostics workflow for photoplethysmography (PPG), inter-beat intervals (IBI), heart-rate variability (HRV) feature preparation, and respiration-proxy displays in `gpbiometrics` projects.

The workflow is diagnostic and signal-processing oriented. The figures show raw and filtered PPG traces, baseline-wander checks, candidate peak markers, IBI sequences, implausible-interval audits, Poincare-style interval displays, segment-level HRV summaries, respiration-proxy traces, and compact dashboard-style reporting. They should not be interpreted as direct evidence of health status, clinical state, diagnosis, stress, emotion, attention, workload, autonomic state, or psychological response.

# Synthetic PPG stream

The example uses synthetic data so that the article is reproducible and does not depend on private recordings.

```{r synthetic-ppg-data}
participants <- sprintf("P%02d", 1:8)
trials <- seq_len(10)
samples_per_trial <- 750
sampling_rate_hz <- 25
time_s <- seq(0, (samples_per_trial - 1) / sampling_rate_hz, length.out = samples_per_trial)

trial_grid <- expand.grid(
  participant_id = participants,
  trial_id = trials,
  KEEP.OUT.ATTRS = FALSE,
  stringsAsFactors = FALSE
)

make_ppg_trial <- function(participant_id, trial_id) {
  condition <- ifelse(trial_id %% 2 == 0, "dense", "simple")
  participant_shift <- match(participant_id, participants) / 120
  heart_hz <- ifelse(condition == "dense", 1.28, 1.18) + participant_shift + rnorm(1, 0, 0.015)
  respiration_hz <- 0.23 + rnorm(1, 0, 0.01)
  baseline <- 0.08 * sin(2 * pi * time_s / 28) + 0.03 * cos(2 * pi * time_s / 13)
  respiratory_modulation <- 1 + 0.18 * sin(2 * pi * respiration_hz * time_s)
  pulse <- respiratory_modulation * sin(2 * pi * heart_hz * time_s)
  harmonic <- 0.25 * sin(2 * pi * heart_hz * 2 * time_s + 0.6)
  noise <- rnorm(length(time_s), 0, ifelse(condition == "dense", 0.045, 0.035))
  ppg_raw <- 1.0 + baseline + 0.32 * pulse + 0.08 * harmonic + noise
  missing <- runif(length(time_s)) < ifelse(condition == "dense", 0.006, 0.003)
  ppg_raw[missing] <- NA_real_
  data.frame(
    participant_id = participant_id,
    trial_id = trial_id,
    condition = condition,
    time_s = time_s,
    ppg_raw = ppg_raw,
    stringsAsFactors = FALSE
  )
}

ppg <- do.call(
  rbind,
  Map(make_ppg_trial, trial_grid$participant_id, trial_grid$trial_id)
)

head(ppg)
```

# Raw and filtered PPG trace

A raw/filtered trace helps document whether preprocessing changes the signal in a plausible and transparent way.

```{r raw-filtered-trace, fig.cap = "Synthetic raw and filtered PPG trace for one participant-trial segment."}
fill_missing <- function(x) {
  if (!anyNA(x)) return(x)
  idx <- seq_along(x)
  observed <- !is.na(x)
  approx(idx[observed], x[observed], xout = idx, rule = 2)$y
}

smooth_series <- function(x, width) {
  y <- stats::filter(x, rep(1 / width, width), sides = 2)
  y <- as.numeric(y)
  idx <- seq_along(y)
  observed <- !is.na(y)
  y[!observed] <- approx(idx[observed], y[observed], xout = idx[!observed], rule = 2)$y
  y
}

one_trial <- ppg[ppg$participant_id == "P01" & ppg$trial_id == 2, ]
one_trial$ppg_filled <- fill_missing(one_trial$ppg_raw)
one_trial$baseline_est <- smooth_series(one_trial$ppg_filled, width = 151)
one_trial$ppg_filtered <- one_trial$ppg_filled - one_trial$baseline_est

plot(
  one_trial$time_s,
  one_trial$ppg_raw,
  type = "l",
  xlab = "Time (s)",
  ylab = "PPG",
  main = "Raw and filtered PPG"
)
lines(one_trial$time_s, one_trial$ppg_filtered + mean(one_trial$ppg_raw, na.rm = TRUE), lty = 2)
legend("topright", legend = c("raw", "filtered, shifted"), lty = c(1, 2), bty = "n")
```

# Baseline-wander check

Baseline-wander displays support transparent reporting of the preprocessing state used for peak detection or interval extraction.

```{r baseline-wander-check, fig.cap = "Synthetic PPG baseline-wander estimate."}
plot(
  one_trial$time_s,
  one_trial$ppg_filled,
  type = "l",
  xlab = "Time (s)",
  ylab = "PPG",
  main = "Baseline-wander check"
)
lines(one_trial$time_s, one_trial$baseline_est, lty = 2)
legend("topright", legend = c("filled raw", "baseline estimate"), lty = c(1, 2), bty = "n")
```

# Candidate PPG peak markers

Candidate peak markers provide a visual audit trail for beat detection. They are algorithmic candidates, not clinical annotations.

```{r ppg-peak-markers, fig.cap = "Synthetic candidate PPG peak markers."}
detect_simple_peaks <- function(y, time_s, threshold = NULL, refractory_s = 0.45) {
  n <- length(y)
  candidate <- which(y[2:(n - 1)] > y[1:(n - 2)] & y[2:(n - 1)] > y[3:n]) + 1
  if (is.null(threshold)) {
    threshold <- median(y, na.rm = TRUE) + 0.35 * stats::sd(y, na.rm = TRUE)
  }
  candidate <- candidate[y[candidate] >= threshold]
  if (length(candidate) == 0) return(integer())
  keep <- candidate[1]
  for (i in candidate[-1]) {
    if ((time_s[i] - time_s[tail(keep, 1)]) >= refractory_s) {
      keep <- c(keep, i)
    }
  }
  keep
}

peak_idx <- detect_simple_peaks(one_trial$ppg_filtered, one_trial$time_s)
peaks <- data.frame(
  time_s = one_trial$time_s[peak_idx],
  ppg_filtered = one_trial$ppg_filtered[peak_idx]
)

plot(
  one_trial$time_s,
  one_trial$ppg_filtered,
  type = "l",
  xlab = "Time (s)",
  ylab = "Filtered PPG",
  main = "Candidate PPG peaks"
)
points(peaks$time_s, peaks$ppg_filtered, pch = 16)
```

# IBI interval sequence

The IBI sequence shows the interval series derived from candidate peak times. Implausible intervals should be flagged before HRV feature extraction.

```{r ibi-sequence, fig.cap = "Synthetic IBI sequence from candidate PPG peaks."}
ibi_ms <- diff(peaks$time_s) * 1000
ibi_data <- data.frame(
  interval_id = seq_along(ibi_ms),
  ibi_ms = ibi_ms
)

if (nrow(ibi_data) >= 12) {
  ibi_data$ibi_ms[5] <- 345
  ibi_data$ibi_ms[12] <- 1620
}

plot(
  ibi_data$interval_id,
  ibi_data$ibi_ms,
  type = "b",
  xlab = "Interval index",
  ylab = "IBI (ms)",
  main = "IBI sequence"
)
abline(h = c(400, 1500), lty = 2)
```

# Implausible interval audit

This audit separates intervals retained for feature extraction from intervals requiring review or correction.

```{r implausible-interval-audit, fig.cap = "Synthetic implausible IBI interval audit."}
ibi_data$flag_implausible <- ibi_data$ibi_ms < 400 | ibi_data$ibi_ms > 1500

plot(
  ibi_data$interval_id,
  ibi_data$ibi_ms,
  type = "n",
  xlab = "Interval index",
  ylab = "IBI (ms)",
  main = "Implausible interval audit"
)
points(ibi_data$interval_id[!ibi_data$flag_implausible], ibi_data$ibi_ms[!ibi_data$flag_implausible], pch = 16)
points(ibi_data$interval_id[ibi_data$flag_implausible], ibi_data$ibi_ms[ibi_data$flag_implausible], pch = 4)
abline(h = c(400, 1500), lty = 2)
legend("topright", legend = c("retained", "flagged"), pch = c(16, 4), bty = "n")
```

# Poincare-style interval display

A Poincare-style interval plot is a compact diagnostic for interval-to-interval structure. It is a descriptive interval display, not a clinical interpretation.

```{r poincare-display, fig.cap = "Synthetic Poincare-style IBI display."}
ibi_clean <- ibi_data$ibi_ms[!ibi_data$flag_implausible]
poincare <- data.frame(
  ibi_n = ibi_clean[-length(ibi_clean)],
  ibi_next = ibi_clean[-1]
)

plot(
  poincare$ibi_n,
  poincare$ibi_next,
  pch = 16,
  xlab = "IBI[n] (ms)",
  ylab = "IBI[n+1] (ms)",
  main = "Poincare-style interval display"
)
abline(0, 1, lty = 2)
```

# Segment-level HRV feature summaries

Segment-level features can be reviewed for coverage and distribution before modelling.

```{r segment-hrv-features, fig.cap = "Synthetic segment-level HRV feature summaries."}
make_hrv_features <- function(participant_id, trial_id) {
  condition <- ifelse(trial_id %% 2 == 0, "dense", "simple")
  mean_ibi <- ifelse(condition == "dense", 780, 835) + rnorm(1, 0, 25)
  sdnn <- ifelse(condition == "dense", 42, 52) + rnorm(1, 0, 6)
  rmssd <- ifelse(condition == "dense", 34, 43) + rnorm(1, 0, 5)
  pnn50 <- pmax(0, pmin(100, ifelse(condition == "dense", 18, 27) + rnorm(1, 0, 5)))
  data.frame(
    participant_id = participant_id,
    trial_id = trial_id,
    condition = condition,
    mean_ibi_ms = mean_ibi,
    sdnn_ms = sdnn,
    rmssd_ms = rmssd,
    pnn50_percent = pnn50,
    stringsAsFactors = FALSE
  )
}

hrv_features <- do.call(
  rbind,
  Map(make_hrv_features, trial_grid$participant_id, trial_grid$trial_id)
)

op <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
boxplot(sdnn_ms ~ condition, data = hrv_features, xlab = "Synthetic condition", ylab = "SDNN (ms)", main = "SDNN summary")
boxplot(rmssd_ms ~ condition, data = hrv_features, xlab = "Synthetic condition", ylab = "RMSSD (ms)", main = "RMSSD summary")
par(op)
```

# Respiration-proxy trace

A respiration-proxy display can be used to document an algorithmic signal estimate derived from PPG. It should be described as a proxy unless validated against a respiration reference.

```{r respiration-proxy, fig.cap = "Synthetic respiration-proxy trace derived from PPG amplitude modulation."}
abs_envelope <- abs(one_trial$ppg_filtered)
resp_proxy <- smooth_series(abs_envelope, width = 101)
resp_proxy <- resp_proxy - mean(resp_proxy, na.rm = TRUE)

plot(
  one_trial$time_s,
  resp_proxy,
  type = "l",
  xlab = "Time (s)",
  ylab = "Respiration proxy",
  main = "PPG-derived respiration proxy"
)
abline(h = 0, lty = 2)
```

# Compact PPG/HRV visual diagnostics dashboard

A compact dashboard can combine preprocessing, peak detection, interval audit, and feature-review displays for technical reporting.

```{r compact-ppg-dashboard, fig.cap = "Compact synthetic PPG/HRV visual diagnostics dashboard."}
op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plot(one_trial$time_s, one_trial$ppg_raw, type = "l", xlab = "Time (s)", ylab = "PPG", main = "Raw PPG")
lines(one_trial$time_s, one_trial$baseline_est, lty = 2)

plot(one_trial$time_s, one_trial$ppg_filtered, type = "l", xlab = "Time (s)", ylab = "Filtered PPG", main = "Candidate peaks")
points(peaks$time_s, peaks$ppg_filtered, pch = 16)

plot(ibi_data$interval_id, ibi_data$ibi_ms, type = "b", xlab = "Interval", ylab = "IBI (ms)", main = "IBI audit")
points(ibi_data$interval_id[ibi_data$flag_implausible], ibi_data$ibi_ms[ibi_data$flag_implausible], pch = 4)
abline(h = c(400, 1500), lty = 2)

boxplot(rmssd_ms ~ condition, data = hrv_features, xlab = "Condition", ylab = "RMSSD (ms)", main = "Feature review")

par(op)
```

# Relation to gpbiometrics helpers

The same reporting pattern can be paired with package-level helpers when project data use the expected column contracts.

```{r helper-map, echo = FALSE}
helper_map <- data.frame(
  task = c(
    "PPG filtering and baseline review",
    "PPG peak detection",
    "IBI quality audit",
    "Beat correction and interval review",
    "HRV feature extraction",
    "Respiration-proxy estimation",
    "Report outputs"
  ),
  representative_helpers = c(
    "filter_gazepoint_ppg_signal(); remove_gazepoint_ppg_baseline_wander()",
    "detect_gazepoint_ppg_peaks(); plot_gazepoint_ppg_peak_detection()",
    "audit_gazepoint_ibi_quality(); filter_gazepoint_ibi_implausible()",
    "correct_gazepoint_beats(); summarize_gazepoint_beat_corrections()",
    "extract_gazepoint_hrv_features(); summarise_gazepoint_hrv_features()",
    "estimate_gazepoint_respiration_from_ppg(); plot_gazepoint_ppg_breathing()",
    "create_gazepoint_qc_supplement(); create_gazepoint_biometrics_report_tables()"
  ),
  stringsAsFactors = FALSE
)

knitr::kable(helper_map)
```

```{r package-helper-pattern, eval = FALSE}
library(gpbiometrics)

ppg_filtered <- filter_gazepoint_ppg_signal(
  biometric_data,
  ppg_col = "PPG",
  time_col = "TIME_MS",
  output_col = "PPG_FILTERED"
)

peaks <- detect_gazepoint_ppg_peaks(
  ppg_filtered,
  ppg_col = "PPG_FILTERED",
  time_col = "TIME_MS",
  participant_col = "participant_id",
  trial_col = "trial_id"
)

ibi_quality <- audit_gazepoint_ibi_quality(
  biometric_data,
  ibi_col = "IBI",
  participant_col = "participant_id",
  trial_col = "trial_id"
)

hrv_features <- extract_gazepoint_hrv_features(
  biometric_data,
  ibi_col = "IBI",
  participant_col = "participant_id",
  trial_col = "trial_id"
)

plot_gazepoint_ppg_peak_detection(ppg_filtered, peaks = peaks)
plot_gazepoint_ppg_poincare(hrv_features)
```

# Reporting recommendation

For manuscripts, technical supplements, and reviewer-facing reports, describe these figures as PPG, IBI, HRV, respiration-proxy, signal-processing, quality-control, and reproducibility diagnostics. Avoid stronger clinical, health-related, affective, workload-related, stress-related, attentional, or psychological interpretation unless supported by validation, controls, and modelling evidence.
