---
title: "Quality-control workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quality-control workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Reviewer-facing QC sequence

This article shows a compact quality-control workflow for Gazepoint-derived biometric and eye-tracking exports. The workflow is conservative: it validates dataset layout, metadata, missingness, signal quality, exclusion decisions, and reproducibility outputs. It does not infer clinical, affective, emotional, diagnostic, or psychological states.

## Dataset layout

```{r}
layout_audit <- check_gazepoint_bids(
  root = "path/to/gazepoint_dataset"
)

layout_audit
```

`check_gazepoint_bids()` is a lightweight Gazepoint-oriented layout audit. It is not a full BIDS validator and does not convert data.

## Metadata and required columns

```{r}
metadata_audit <- validate_gazepoint_metadata(
  data = gazepoint_data,
  required_columns = c("participant_id", "time", "trial")
)

metadata_audit
```

## Missingness

```{r}
missingness_summary <- summarize_gazepoint_missingness(gazepoint_data)
missingness_summary

plot_gazepoint_missingness(gazepoint_data)
```

Missingness summaries and plots are descriptive QC tools. They help identify sparse recordings, dropouts, and participant-level recording problems.

## Pupil and signal artifacts

```{r}
pupil_qc <- detect_gazepoint_blinks(
  gazepoint_data,
  pupil_cols = c("pupil_left", "pupil_right")
)

nonwear_qc <- detect_gazepoint_nonwear(
  gazepoint_data,
  signal_cols = c("eda", "ppg", "ecg")
)
```

Masking, interpolation, filtering, and beat correction should be reported as transparent preprocessing steps.

## Signal quality and rule-based flags

```{r}
signal_quality <- compute_gazepoint_signal_quality(
  gazepoint_data,
  signal_cols = c("eda", "ppg", "ecg")
)

quality_summary <- summarize_gazepoint_signal_quality(signal_quality)

quality_flags <- classify_gazepoint_signal_quality(
  signal_quality,
  rules = list(
    missingness_rate = 0.20,
    quality_index = 0.50
  )
)
```

Rule-based quality flags support auditability. They are not clinical, psychological, or affective interpretations.

## Exclusion recommendations

```{r}
exclusion_recommendations <- recommend_gazepoint_biometric_exclusions(
  quality_flags
)

exclusion_recommendations
```

Exclusion helpers summarize evidence for review. They support analyst judgment but do not replace it.

## Static dashboard

```{r}
dashboard <- pipeline_comparison_dashboard(
  quality_flags,
  participant_col = "participant_id",
  session_col = "session",
  missingness_col = "missingness_rate",
  quality_col = "quality_index",
  qc_status_col = "qc_status",
  failed_rules_col = "failed_rules",
  excluded_col = "excluded",
  notes_col = "audit_notes"
)

dashboard
```

`pipeline_comparison_dashboard()` creates a static reviewer-facing summary. It is not a Shiny app, GUI dashboard, model, or interpretation layer.

## Reproducibility outputs

```{r}
manifest <- create_gazepoint_analysis_manifest(
  input_files = list.files("path/to/gazepoint_dataset", recursive = TRUE),
  parameters = list(
    missingness_threshold = 0.20,
    quality_threshold = 0.50
  )
)

dictionary <- create_gazepoint_dictionary(gazepoint_data)
```

## Recommended reporting language

> Gazepoint-derived exports were processed using conservative, rule-based quality-control checks. Dataset layout, metadata fields, missingness, signal quality, artifact flags, and exclusion recommendations were summarized before analysis. The workflow produced audit-ready summaries and reproducibility metadata, but did not infer clinical, affective, emotional, or diagnostic states from biometric signals.

## Summary

The recommended sequence is: layout check, metadata validation, missingness review, artifact detection, signal-quality scoring, rule-based flags, exclusion review, dashboard summary, and reproducibility output.
