---
title: "Visual QC dashboard workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visual QC dashboard workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.6,
  dpi = 96
)
set.seed(20260710)
```

# Purpose

This article demonstrates a compact visual quality-control dashboard workflow for `gpbiometrics` projects.

The goal is to show how synthetic Gazepoint-like biometric streams can be summarized visually before modelling or reporting. The plots are diagnostic and reproducibility-oriented. They should not be interpreted as direct evidence of emotion, stress, attention, workload, health status, clinical state, diagnosis, or psychological response.

# Synthetic biometric stream

The example uses synthetic data so that the article is reproducible and does not depend on private recordings.

```{r synthetic-data}
participants <- sprintf("P%02d", 1:10)
trials <- seq_len(12)
samples_per_trial <- 120

dat <- expand.grid(
  participant_id = participants,
  trial_id = trials,
  sample_id = seq_len(samples_per_trial),
  KEEP.OUT.ATTRS = FALSE,
  stringsAsFactors = FALSE
)

dat$time_ms <- (dat$sample_id - 1) * 16.67
dat$condition <- ifelse(dat$trial_id %% 2 == 0, "dense", "simple")
dat$session <- ifelse(match(dat$participant_id, participants) <= 5, "morning", "afternoon")

participant_offset <- match(dat$participant_id, participants) / 30
trial_offset <- dat$trial_id / 80
phase <- dat$sample_id / samples_per_trial

dat$pupil_mm <- 3.1 + participant_offset + 0.15 * sin(2 * pi * phase) + rnorm(nrow(dat), 0, 0.05)
dat$gsr_us <- 1.4 + trial_offset + 0.08 * cos(2 * pi * phase) + rnorm(nrow(dat), 0, 0.03)
dat$ppg <- 0.8 + 0.25 * sin(2 * pi * phase * 8) + rnorm(nrow(dat), 0, 0.04)
dat$hr_bpm <- 72 + 3 * sin(2 * pi * phase) + participant_offset * 8 + rnorm(nrow(dat), 0, 1.2)

p_miss <- 0.015 + 0.008 * (dat$condition == "dense") + 0.006 * (dat$session == "afternoon")
g_miss <- 0.010 + 0.006 * (dat$trial_id > 8)
p_bad <- runif(nrow(dat)) < p_miss
g_bad <- runif(nrow(dat)) < g_miss

dat$pupil_mm[p_bad] <- NA_real_
dat$gsr_us[g_bad] <- NA_real_

head(dat)
```

# Channel-level missingness

The first plot summarizes sample-level missingness by measurement stream. This is a coverage check, not a substantive result.

```{r channel-missingness, fig.cap = "Synthetic channel-level missingness summary."}
channel_missing <- data.frame(
  channel = c("Pupil", "GSR", "PPG", "HR"),
  missing_percent = c(
    mean(is.na(dat$pupil_mm)) * 100,
    mean(is.na(dat$gsr_us)) * 100,
    mean(is.na(dat$ppg)) * 100,
    mean(is.na(dat$hr_bpm)) * 100
  )
)

barplot(
  channel_missing$missing_percent,
  names.arg = channel_missing$channel,
  ylim = c(0, max(5, channel_missing$missing_percent) * 1.25),
  ylab = "Missing samples (%)",
  xlab = "Measurement stream",
  main = "Missingness by stream"
)
```

# Participant-by-trial coverage

A participant-by-trial heatmap helps identify concentrated gaps that may not be visible from a global percentage.

```{r participant-trial-coverage, fig.cap = "Synthetic participant-by-trial pupil missingness heatmap."}
pt_missing <- aggregate(
  is.na(dat$pupil_mm),
  by = list(participant_id = dat$participant_id, trial_id = dat$trial_id),
  FUN = mean
)
names(pt_missing)[3] <- "pupil_missing_rate"

miss_matrix <- xtabs(pupil_missing_rate * 100 ~ participant_id + trial_id, pt_missing)

image(
  x = as.integer(colnames(miss_matrix)),
  y = seq_len(nrow(miss_matrix)),
  z = t(miss_matrix[nrow(miss_matrix):1, ]),
  axes = FALSE,
  xlab = "Trial",
  ylab = "Participant",
  main = "Pupil missingness by participant and trial"
)
axis(1, at = as.integer(colnames(miss_matrix)), labels = colnames(miss_matrix))
axis(2, at = seq_len(nrow(miss_matrix)), labels = rev(rownames(miss_matrix)), las = 1)
box()
```

# Signal activity over time

Time-course displays are useful for checking whether streams are present and whether the plotted signal has plausible continuity in the exported file.

```{r signal-activity, fig.cap = "Synthetic signal activity for one participant-trial segment."}
one_trial <- dat[dat$participant_id == "P01" & dat$trial_id == 1, ]

op <- par(mfrow = c(3, 1), mar = c(3.5, 4, 2, 1))
plot(one_trial$time_ms, one_trial$pupil_mm, type = "l", xlab = "Time (ms)", ylab = "Pupil (mm)", main = "Pupil stream")
plot(one_trial$time_ms, one_trial$gsr_us, type = "l", xlab = "Time (ms)", ylab = "GSR (uS)", main = "GSR stream")
plot(one_trial$time_ms, one_trial$ppg, type = "l", xlab = "Time (ms)", ylab = "PPG", main = "PPG stream")
par(op)
```

# Trial-level quality score

A simple quality score can help order trial-level review. The score below is only a synthetic demonstration of a QC index.

```{r quality-score, fig.cap = "Synthetic trial-level QC score by condition."}
pupil_miss <- aggregate(is.na(dat$pupil_mm), by = list(participant_id = dat$participant_id, trial_id = dat$trial_id, condition = dat$condition), FUN = mean)
names(pupil_miss)[4] <- "pupil_missing"

gsr_miss <- aggregate(is.na(dat$gsr_us), by = list(participant_id = dat$participant_id, trial_id = dat$trial_id), FUN = mean)
names(gsr_miss)[3] <- "gsr_missing"

qc <- merge(pupil_miss, gsr_miss, by = c("participant_id", "trial_id"))
qc$qc_score <- pmax(0, 100 - 100 * qc$pupil_missing - 80 * qc$gsr_missing)

boxplot(
  qc_score ~ condition,
  data = qc,
  ylim = c(0, 105),
  ylab = "QC score",
  xlab = "Synthetic condition",
  main = "Trial-level QC score"
)
stripchart(qc_score ~ condition, data = qc, vertical = TRUE, method = "jitter", pch = 16, add = TRUE)
```

# Compact QC dashboard

A dashboard view can combine missingness, trial-level scores, timing coverage, and selected stream traces into one report-oriented display.

```{r compact-dashboard, fig.cap = "Compact synthetic visual QC dashboard."}
trial_counts <- aggregate(sample_id ~ participant_id + trial_id, dat, length)
names(trial_counts)[3] <- "n_samples"

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

barplot(
  channel_missing$missing_percent,
  names.arg = channel_missing$channel,
  ylab = "Missing (%)",
  main = "Missingness"
)

boxplot(
  qc$qc_score,
  ylab = "QC score",
  main = "Trial QC scores"
)

hist(
  trial_counts$n_samples,
  xlab = "Samples per participant-trial",
  main = "Sample coverage",
  breaks = 10
)

plot(
  one_trial$time_ms,
  one_trial$ppg,
  type = "l",
  xlab = "Time (ms)",
  ylab = "PPG",
  main = "Example PPG trace"
)

par(op)
```

# Relation to gpbiometrics helpers

The same reporting pattern can be paired with package-level helpers when the project data use the expected column contracts.

```{r helper-map, echo = FALSE}
helper_map <- data.frame(
  task = c(
    "Missingness review",
    "Signal-quality review",
    "Activity and timebase review",
    "Dashboard reporting",
    "Decision logging"
  ),
  representative_helpers = c(
    "summarize_gazepoint_missingness(); plot_gazepoint_missingness()",
    "summarize_gazepoint_signal_quality(); plot_gazepoint_signal_quality()",
    "audit_gazepoint_signal_activity(); audit_gazepoint_time_resets()",
    "create_gazepoint_quality_dashboard(); plot_gazepoint_biometric_report_dashboard()",
    "create_gazepoint_analysis_decision_log(); write_gazepoint_decision_log()"
  ),
  stringsAsFactors = FALSE
)

knitr::kable(helper_map)
```

```{r package-helper-pattern, eval = FALSE}
library(gpbiometrics)

missingness <- summarize_gazepoint_missingness(
  biometric_data,
  group_cols = c("participant_id", "trial_id")
)

plot_gazepoint_missingness(missingness)

signal_quality <- summarize_gazepoint_signal_quality(
  biometric_data,
  participant_col = "participant_id",
  trial_col = "trial_id"
)

plot_gazepoint_signal_quality(signal_quality)

activity <- audit_gazepoint_signal_activity(
  biometric_data,
  time_col = "TIME_MS",
  participant_col = "participant_id",
  trial_col = "trial_id"
)

plot_gazepoint_signal_activity(activity)
```

# Reporting recommendation

For manuscripts, technical supplements, and reviewer-facing reports, describe these figures as QC, timing, coverage, signal-processing, or reproducibility displays. Stronger substantive interpretation should be reserved for designs with appropriate validation, controls, and modelling evidence.
